Fast, Scalable Phrase-Based SMT Decoding

### 2016-01-29 acl.2016/acl2016.tex
acl paper

SMT has been one of the outstanding success story from the NLP community in the last decade. It has transition from a mostly research discipline to services such as Google Translate, Microsoft Translator Hub, as well as services and products built around offline products such as the open-source Moses toolkit. The latter has spawned a cottage industry  encompassing a range of organizations and services from small language service providers that use SMT to reduce translation cost to large inter-govermental organizations such as the EU and the UN that provides high volume translation.

For high volume users, decoding is a largest and most critical part of the translation process which needs to be fast and efficient. However, it has been noticed that the Moses decoder, amongst others, is unable to efficiently use multiple CPU cores that are now common on modern servers (reviewed paper, github discussion). That is, the time taken to decode a test set does not substantial decrease when more cores are used, in fact, decoding time may increase when more cores are added. The issue will only become more noticeable as the commercial use of SMT grows and the number of cores in servers increases.

There have be speculation on the causes of the inefficiency as well as remedies. This paper is the first we know of that seeks to tackle this problem head on. We present an phrase-based decoder that is not only significantly faster than the Moses baseline for single-threaded operation, but is able to scale run multiple threads on multicore machines with only a slightly loss in linear speed. Model scores and functionality are compatible with Moses to aid comparison and ease of transition for users. All source code will be made available under an open-source license.

>>>>

SMT has been one of the outstanding success story from the NLP community in the last decade, progressing from a mostly research discipline to public useability via services such as Google Translate, Microsoft Translator Hub, as well as services and products built around offline products such as Language Weaver and the open-source Moses toolkit. The latter has spawned a cottage industry  encompassing a range of organizations and services from small language service providers seeking to reduce translation cost, to large inter-govermental organizations such as the EU and the UN that require high volume, high quality translation.

For high volume users, decoding is a largest and most critical part of the translation process which needs to be fast and efficient. However, it has been noticed that the Moses decoder, amongst others, is unable to efficiently use multiple CPU cores that are now common on modern servers (reviewed paper, github discussion). That is, the time taken to decode a test set does not substantial decrease when more cores are used, in fact, decoding time may increase when more cores are added. The problem will continue to grow as the commercial use of SMT increases and the number of CPU cores increases.

There have be speculation on the causes of the inefficiency as well as potential remedies. This paper is the first we know of that seeks to tackle this problem head on. We present an phrase-based decoder that is not only significantly faster than the Moses baseline for single-threaded operation, but is able to scale run multiple threads on multicore machines with only a slightly loss in linear speed. Model scores and functionality are compatible with Moses to aid comparison and ease of transition for users. All source code will be made available under an open-source license.
